---
title: "Cazelles et al. (2016)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cazelles_2016}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biogeonet)
```


## Philosophy

In Cazelles[^ref], we basically expand the classical Theory of Island Biogeography[^ref2] to include ecological interactions and a link to the environmental niche. To do so we:

1. simulated ecological networks of 10 species using the niche model[^ref3]
2. link the simulated ecological interactions to extinction probability
3. link an arbitrary environmental gradient to the colonization probability
4. use all the steps above to create the transition matrix of a Markov chain (from any of the $2^{10}$ communities to any other) and then solve it to obtain the presence absence of all communities.


## Niche model

To simulate I use the niche model that I implemented in `nicheModel()`. The first parameters, `nbsp`, determines the number of species:

```{r}
nicheModel(nbsp = 5, .2, 1)
nicheModel(nbsp = 10, .2, 1)
```

The second parameter controls the connectance (the expected connectance).

```{r}
nicheModel(5, connec = .1, 1)
nicheModel(5, connec = .4, 1)
```

The last parameter, `mode`, affects the sign structure of the output matrix (not included in the original publication of the niche model):

```{r}
# mode 1 + and -, refered as "predation" in Cazelles (2016)
nicheModel(5, .2, mode = 1)
# mode 2, only -, refered as "competition" in Cazelles (2016)
nicheModel(5, .2, mode = 2)
# mode 2, only +, refered as "mutualism" in Cazelles (2016)
nicheModel(5, .2, mode = 3)
```

See `?nicheModel` for further details.


## Colonization

The colonisation probability is a bell-shaped function (see `?colonization` for details).


```{r}
seqx <- seq(0, 10, .01)
vc_colo <- sapply(seqx, FUN = colonization, 1, 5, 2)
plot(seqx, vc_colo, type="l", las = 1, ylab = "Extinction probability",
  xlab = "Environmental gradient")
```

## Extinction

The extinction probability is basically a flexible logistic function (see `?extinction` for details)

```{r}
seqx <- seq(-2, 2, .01)
vc_exti <- sapply(seqx, FUN = extinction, 1e-5, 1e-3, 1e3, -1)
plot(seqx, log10(vc_exti), type="l", las = 1, ylab = "Extinction probability",
  xlab = "Environmental gradient")
```


## Building Markov chain and deriving its equilibrium

For a specific vector of colonisation (1 value per species), a specific
ecological networks (use to derive extinction probabilities), we can build the
transition matrix of a Markov chain with `buildMarkov()`. So far, it builds the matrix of a continuous-time Markov chain](https://en.wikipedia.org/wiki/Markov_chain#Continuous-time_Markov_chain) ans so probabilities described above should actuall be rates.

```{r}
# number of species
nbsp <- 4
# network
web <- nicheModel(nbsp, .2, mode = 1)
# colonisation vector
colo <- 10^runif(nbsp, -7, -3)
#
mat <- buildMarkov(nbsp, colo, web, 1e-5, 1e-3, 1e3, -1)
mat
```

then `solveMarkov()` returns the probabilities at equilibrium for all communities:

```{r}
sol <- solveMarkov(mat, continuous = TRUE)
sol
```

This gives the probability at equilibrium (*sensu* MacArthur & Wilson 1967)  for all of the $2^{nbsp}$ community (here `nbsp`=4, so there are 16 potential communities).

```{r}
# NB some to one
sum(sol)
```



## Reproduce the study

In Cazelles *et al.* (2016), we basically solved tons of those for 10 species
(1024 communities), for the 3 sign structure described above and under various
scenarios of connectance and environmental conditions (so a bunch of for loops using the steps described above!).




[^ref]: Cazelles, K., Mouquet, N., Mouillot, D. & Gravel, D. 2016. On the integration of biotic interaction and environmental constraints at the biogeographical scale. Ecography. 39, 921–931. [DOI: 10.1111/ecog.01714](http://onlinelibrary.wiley.com/doi/10.1111/ecog.01714/abstract).

[^ref2]: MacArthur, R. H. and Wilson, E. O. 1967. Theory of island
biogeography. – Princeton Univ. Press. [ISBN: 0691088365](https://press.princeton.edu/books/paperback/9780691088365/the-theory-of-island-biogeography).

[^ref3]: Williams, R. J. and Martinez, N. D. 2000. Simple rules yield
complex food webs. – Nature 404: 180–183.[DOI:10.1038/35004572](https://doi.org/10.1038/35004572).
